---
title: 'Math Final Project'
author: "Judy Jiang, Aitana Hierro,Virti Sheth, Agustin Arcos"

date: "Dec.3.2023"
output:
  html_document: default
  header-includes: \usepackage{color}
fontsize: 12pt
margin: 1in
---
```{r setup, echo=FALSE, message=F, warning=F}

#==============================================================================
#   This chunk will be used for every assignment
#==============================================================================

library(alabama)
library(numDeriv)
```



```{r Question1 , echo=FALSE, message=FALSE, warning=FALSE, include=FALSE}

### Define the cost function
f <- function(theta) {
  x <- theta[1:5]
  y <- theta[6:10]
  z <- theta[11:12]
  return(60*sum(x)+60*sum(y)+60*(theta[1]+theta[2]+theta[3]+theta[6]+theta[7]+theta[8]+theta[11])+90*(theta[4]+theta[5]+theta[9]+theta[10]+theta[11]+2*theta[12]))
}


### Define the constrained inequality functions
g <- function(theta) {
  h <- 0
  h[1] <- 6*theta[1]-32
  h[2] <- 6*theta[6]-8
  h[3] <- 6*theta[2]-68
  h[4] <- 6*theta[7]-17
  h[5] <- 6*(theta[1] + theta[3])-56
  h[6] <- 6*(theta[6] + theta[8])-14
  h[7] <- 6*(theta[2] + theta[4])-76
  h[8] <- 6*(theta[7] + theta[9])-19
  h[9] <- 6*(theta[3] + theta[5] + theta[11])-64
  h[10] <- 6*(theta[8] + theta[12])-16
  h[11] <- 6*(theta[4] + theta[11] + theta[12])-28
  h[12] <- 6*theta[9]-7
  h[13] <- 6*(theta[5] + theta[12])-8
  h[14] <- 6*theta[10]-2
  return(h)
}


## initial values
p0 <- c(20,20,20,20,20,20,20,20,20,20,20,20)


### Optimize the function
k <- constrOptim.nl(p0, f, hin = g)
# Round up the parameter values in k$par
k$par <- ceiling(k$par)
#theta = k$par

### print the value of each variable and minimize costs
print(k$par)
print(f(k$par))


```
a) How many full-time English-speaking agents, full-time Spanish-speaking agents, and part-time agents should Alex hire for each 2-hour shift to minimize operating costs while attending to all calls?

b) What is the minimum cost for the optimization model to assist Alex’s decision in hiring all agents that she needs?

```{r A1 , echo=FALSE, message=FALSE, warning=FALSE}


# Assuming the following counts of agents for each category
full_time_english_agents <- c(6, 12, 7, 1, 1, 0, 0)  # Replace with your actual values
full_time_spanish_agents <- c(2, 3, 3, 2, 1, 0, 0)  # Replace with your actual values
part_time_english_agents <- c(0, 0, 0, 0, 0, 4, 1)  # Replace with your actual values

# Create a data frame with the desired information
agents_table <- data.frame(
  'TIME SLOT' = c("7-9 am", "9-11 am", "11-13 pm", "13-15 pm", "15-17 pm", "17-19 pm", "19-21 pm"),
  'FT.ENGLISH' = full_time_english_agents,
  'FT.SPANISH' = full_time_spanish_agents,
  'ST.ENGLISH' = part_time_english_agents
)



# Print the table
print(agents_table)

print(f(k$par))
```



```{r Q2 , echo=FALSE, message=FALSE, warning=FALSE, include=FALSE}

## Question c and d
f <- function(theta) {
  x <- theta[1:5]
  y <- theta[6:10]
  z <- theta[11:12]
  return(60*sum(x)+60*sum(y)+60*(theta[1]+theta[2]+theta[3]+theta[6]+theta[7]+theta[8]+theta[11])+90*(theta[4]+theta[5]+theta[9]+theta[10]+theta[11]+2*theta[12]))
}


### Define the constrained inequality functions, with constrain each variable should greater than 0
g <- function(theta) {
  h <- 0
  h[1] <- 6*theta[1]-32
  h[2] <- 6*theta[6]-8
  h[3] <- 6*theta[2]-68
  h[4] <- 6*theta[7]-17
  h[5] <- 6*(theta[1] + theta[3])-56
  h[6] <- 6*(theta[6] + theta[8])-14
  h[7] <- 6*(theta[2] + theta[4])-76
  h[8] <- 6*(theta[7] + theta[9])-19
  h[9] <- 6*(theta[3] + theta[5] + theta[11])-64
  h[10] <- 6*(theta[8] + theta[12])-16
  h[11] <- 6*(theta[4] + theta[11] + theta[12])-28
  h[12] <- 6*theta[9]-7
  h[13] <- 6*(theta[5] + theta[12])-8
  h[14] <- 6*theta[10]-2
  return(h)
}

p0 <- c(20,20,20,20,20,20,20,20,20,20,20,20)


### Define equality constrain (new constrain)
g2 <- function(theta){
  h=0
  h[1]=theta[4]-1
  h[2]=theta[5]-1
  return(h)
}

### Optimize the function
k <- constrOptim.nl(p0, f, hin = g, heq = g2)
# Round up the parameter values in k$par
k$par <- ceiling(k$par)
f(k$par)


print(k$par)
print(f(k$par))
```


c) How many full-time English-speaking agents, full-time Spanish-speaking agents, and part-time agents should Alex hire for each 2-hour shift to minimize operating costs while attending to all calls? 

d) What is the minimum cost for the optimization model to assist Alex’s decision in hiring all agents that she needs?

```{r A2 , echo=FALSE, message=F, warning=FALSE}

# Assuming the following counts of agents for each category
full_time_english_agents <- c(6, 13, 7, 1, 1, 0, 0)  # Replace with your actual values
full_time_spanish_agents <- c(2, 3, 3, 2, 1, 0, 0)  # Replace with your actual values
part_time_english_agents <- c(0, 0, 0, 0, 0, 4, 1)  # Replace with your actual values

# Create a data frame with the desired information
agents_table <- data.frame(
  'TIME SLOT' = c("7-9 am", "9-11 am", "11-13 pm", "13-15 pm", "15-17 pm", "17-19 pm", "19-21 pm"),
  'FT.ENGLISH' = full_time_english_agents,
  'FT.SPANISH' = full_time_spanish_agents,
  'ST.ENGLISH' = part_time_english_agents
)

# Print the table
print(agents_table)
print(f(k$par))

```

```{r Question3 , echo=FALSE, message=FALSE, warning=FALSE, include=FALSE}
# Question e & f

f <- function(theta) {
  x <- theta[1:5]
  z <- theta[6:7]
  return(60*sum(x)+60*(theta[1]+theta[2]+theta[3]+theta[6])+90*(theta[4]+theta[5]+theta[6]+2*theta[7]))
}


### Define the constrained inequality functions
g <- function(theta) {
  h <- 0
  h[1] <- 6*theta[1]-40
  h[2] <- 6*theta[2]-85
  h[3] <- 6*(theta[1]+theta[3])-70
  h[4] <- 6*(theta[2]+theta[4])-95
  h[5] <- 6*(theta[3] + theta[5]+theta[6])-80
  h[6] <- 6*(theta[4] + theta[6]+theta[7])-35
  h[7] <- 6*(theta[5] + theta[7])-10
  h[8] <- theta[1]
  h[9] <- theta[2]
  h[10] <- theta[3]
  h[11] <- theta[4]
  h[12] <- theta[5]
  h[13] <- theta[6]
  h[14] <- theta[7]
  return(h)
}

p0 <- c(20,20,20,20,20,20,20)


### Optimize the function
k <- constrOptim.nl(p0, f, hin=g)
k$par <- ceiling(k$par)
k$value <- round(k$value)

print(k$par)
print(f(k$par))


# Assuming the following counts of agents for each category
full_time_agents <- c(7, 15, 8, 2, 2, 0, 0)  # Replace with your actual values
part_time_agents <- c(0, 0, 0, 0, 5, 1, 0)  # Replace with your actual values

# Create a data frame with the desired information
agents_table <- data.frame(
  'TIME SLOT' = c("7-9 am", "9-11 am", "11-13 pm", "13-15 pm", "15-17 pm", "17-19 pm", "19-21 pm"),
  'FULL TIME AGENTS' = full_time_english_agents,
  'PART TIME AGENTS' = part_time_english_agents
)

# Print the table
print(agents_table)

```

e) How many full-time and part-time agents should Alex hire for each 2-hour shift to
minimize operating costs while attending to all calls? 

f) What is the minimum cost for the optimization model to assist Alex’s decision in
hiring all agents that she needs? 



```{r A3 , echo=FALSE, message=FALSE, warning=FALSE}


# Assuming the following counts of agents for each category
full_time_agents <- c(7, 15, 8, 2, 2, 0, 0)  # Replace with your actual values
part_time_agents <- c(0, 0, 0, 0, 5, 1, 0)  # Replace with your actual values

# Create a data frame with the desired information
agents_table <- data.frame(
  'TIME SLOT' = c("7-9 am", "9-11 am", "11-13 pm", "13-15 pm", "15-17 pm", "17-19 pm", "19-21 pm"),
  'FULL TIME AGENTS' = full_time_english_agents,
  'PART TIME AGENTS' = part_time_english_agents
)

# Print the table
print(agents_table)
print(f(k$par))

```


g) What is the maximum percentage increase in the hourly wage rate that Alex can offer
to bilingual agents over monolingual agents without increasing the total operating
costs? 


```{r Question4, echo=FALSE, message=FALSE, warning=FALSE}
rate <- ((5610 - 5130) / 5610)
percentage <- round(rate * 100, 1)
print(sprintf("%s%%", percentage))

```

  



